---
title: "Synaptosomal vs Cytosolic"
output: pdf_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F, message = F, warning = F
)
```

```{r setup}
library(KRSA)
library(knitr)
library(tidyverse)
library(furrr)
```

```{r model, warning=F,include=T, cache=TRUE}

# read processed data

readRDS("../processed_data/global/data_modelded.rds") -> data_modeled
readRDS("../processed_data/global/qc_peptides.rds") -> new_pep
readRDS("../processed_data/global/data_pw_200.rds") -> data_pw_200

```


## Group Comparison

To compare between samples, a two-group comparison is performed. In this case, the two group comparisons are:

* SYNAPTOSOMAL vs CYTOSOLIC  
 
The *Slope_Transforemed* ratio between each group, paired by chip, is calculated to the fold change. Based on the fold change, peptides that pass a certain fold change threshold are considered significant hits. Also, quality control steps applied in each comparison to filter out peptides that do not reach specific criteria:

-   The *Medain_SigmBg* at max exposure *200ms* must be above a certain value\
-   $R^2$ of the linear model fit must be above a threshold value

These *Filtering Parameters* (fold change threshold, QC criteria) can be modified to adjust the stringency of the analysis. The main *Filtering Parameters* that are used for this analysis:

-   The *Medain_SigmBg* at max exposure *200ms* must be equal or above 5\
-   $R^2$ of the linear model fit must be above or equal 0.9\
-   Log fold change (LFC) cutoffs at (0.6, 1.1, 1.6, 2.1, 2.6)

```{r groupDiff1,echo=F, fig.height=8,fig.width=8,fig.align="center", cache=TRUE}

# Define Groups to be compared
comparisons <- list(Comp1 = c("SYNAPTOSOMAL", "CYTOSOLIC"))

# This function will run both QC steps (krsa_filter_lowPeps, krsa_filter_nonLinear) and krsa_filter_ref_pep
krsa_quick_filter(data = data_pw_200, data2 = data_modeled$scaled,
    signal_threshold = 5, r2_threshold = 0.9,
    groups = comparisons$Comp1) -> pep_passed_qc

# This function calculates log2 fold change values between the defined groups
# The byChip argument lets you calculates the log2 fold change the results within each chip
krsa_group_diff(data_modeled$scaled, comparisons$Comp1, pep_passed_qc, byChip = T) -> diff_df

#sve LFC table
#write_delim(diff_df, "SYNAPTOSOMAL_vs_CYTOSOLIC__LFC_table.txt", delim = "\t")


# Extract top peptides based on the LFC cutoff
krsa_get_diff(diff_df,totalMeanLFC ,c(0.6, 1.1, 1.6, 2.1, 2.6)) %>% list("meanLFC" = .) -> sigPeps
krsa_get_diff_byChip(diff_df, LFC, c(0.6, 1.1, 1.6, 2.1, 2.6)) -> sigPepsPerChip

# Combine the peptides hits in one list
sigPeps_total <- list(sigPeps, sigPepsPerChip) %>% unlist(recursive = F) %>%  unlist(recursive = F)


```

\newpage



### SYNAPTOSOMAL vs CYTOSOLIC (Upstream Kinase Analysis)

The lab carefully curated and mapped the kinases that can act and phosphorylate each peptide present on the chip. This was achieved by using multiple sources including GPS 3.0, Kinexus Phosphonet, PhosphoELM and PhosphoSite Plus. Based on that association between peptides and kinases, a random sampling analysis is performed for these hits. The basic idea of *KRSA* is: For each iteration (*2000* iterations performed in this analysis), the same number of hits are randomly selected from the total 141/or 193 peptides present on the chip. Predicted kinases are then mapped to this sample list of peptides and number of kinases are determined. The kinase count from the actual hits and random sampling is then compared to determine the significance.

```{r krsa1, echo=F, warning=F, message=FALSE, fig.align="center", fig.height=8, fig.width=8, cache=TRUE}


# load in chip coverage and kinase-substrate files OR upload your own files
# if PTK chip, use:
# chipCov <- KRSA_coverage_PTK_PamChip_86402_v1
# KRSA_file <- KRSA_Mapping_PTK_PamChip_86402_v1

# STK chip
chipCov <- KRSA_coverage_STK_PamChip_87102_v2
KRSA_file <- KRSA_Mapping_STK_PamChip_87102_v1



# run the KRSA function to do the random sampling analysis, set seed that can be used later to reproduce results, and choose number of iterations
krsa(sigPeps$meanLFC$`0.6`, return_count = T, seed = 123, itr = 2000,
     map_file = KRSA_file, cov_file = chipCov) -> fin

# View the Z score table
kable(head(fin$KRSA_Table,25), digits = 3)


# find top and bottom kinases
bothways <- c(pull(head(fin$KRSA_Table, 10), Kinase), pull(tail(fin$KRSA_Table, 10), Kinase))

# Use these kinase to generate histogram plots for each selected kinase
krsa_histogram_plot(fin$KRSA_Table, fin$count_mtx, bothways)


# For parallel computing, load the furrr package:
# opens multiple R sessions to run faster
plan(multisession)

# Run the KRSA function across the different sets of peptides using the furrr package for parallel computing
future_map(sigPeps_total, krsa) -> mutiple_krsa_outputs


# For none parallel computing:
# Run KRSA function across the different sets of peptides
#map(sigPeps_total, krsa) -> mutiple_krsa_outputs

# Tidy output
df <- data.frame(matrix(unlist(mutiple_krsa_outputs), ncol = max(lengths(mutiple_krsa_outputs)), byrow = TRUE))
df <- setNames(do.call(rbind.data.frame, mutiple_krsa_outputs), names(mutiple_krsa_outputs$meanLFC.0.6))

df %>% rownames_to_column("method") %>% select(Kinase, Z, method) %>% 
  mutate(method = str_extract(method, "\\w+\\.\\w+\\.\\w+")) %>% 
  mutate(method = gsub("(^\\w+)[\\.]", "\\1>", method)) %>% 
  mutate_if(is.numeric, round, 2) -> df

df %>% 
  pivot_wider(names_from = method, values_from = Z) -> df2
  

# Creates an average Z score table using the across chip analysis
df %>% 
  filter(grepl("mean", method)) %>% 
  select(Kinase, Z, method) %>% group_by(Kinase) %>% mutate(AvgZ = mean(Z)) -> AvgZTable

# save file
AvgZTable %>% write_delim("../processed_data/comparisons/SYNAPTOSOMAL_vs_CYTOSOLIC_KRSA_Table_acrossChip.txt", delim = "\t")

# Creates an average Z score table using the within chip analysis
df %>% 
  filter(!grepl("mean", method)) %>% 
  select(Kinase, Z, method) %>% group_by(Kinase) %>% mutate(AvgZ = mean(Z)) -> AvgZTable2

# save file
AvgZTable2 %>% write_delim("../processed_data/comparisons/SYNAPTOSOMAL_vs_CYTOSOLIC_KRSA_Table_withinChip.txt", delim = "\t")


```
